### 2024-04-23 Paper/paper.tex
Updates from Overleaf

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
\usepackage{natbib}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{papercitations.bib}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}

@@

	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from the Hospital Readmissions Reduction Program}

@@

    \begin{abstract}
		{\small
		The objectives that underlie nonprofit firm behaviors are not well understood, but perhaps that is because there are characteristics even within nonprofits that dictate behavior. In this paper, I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether having clinical experience on a hospital executive team affects hospital behavior. In particular, I leverage the Hospital Readmissions Reduction Program of 2012 as a shock to similar quality hospitals, and estimate the difference in response from hospitals with and without clinical experience. I find that while all penalized hospitals decreased readmissions after the shock, non-clinical executive team hospital decreased readmissions at a faster rate than those with a clinical executive team. The difference equates to approximately 15 additional readmissions for clinical team hospitals. I find no differential effect of clinical executive teams on mortality rates.

@@

  The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms. Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing mission-based purposes. Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity tradeoff (Steinberg, 1986). An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (\cite{sloan2000not}). In this paper, I leverage a large-scale shock to nonprofit hospitals in the US to investigate whether characteristics of the objective function are revealed by the composition of their leadership team.

  Hospitals in particular have been a common focus when seeking to understand not-for-profit firms. This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits. Second, nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}). Thus, nonprofit behavior in the hospital industry is relevant and directly affects consumers of health care. However, it remains unclear what type of objective function drives nonprofit hospitals (\cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).

  What has not been considered is how the leadership team within nonprofits potentially drives objectives. In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than just for-profit or not-for-profit status. Within nonprofit hospitals, there is variation in the propensity to hire executives with a background in clinical work, which could reasonably change how much weight the hospital places on profit vs. quality of care in their objective function. I contribute to our understanding of hospital and nonprofit behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.

  Using publicly available tax forms, I construct a novel data set of nonprofit hospitals from 2009-2015 which captures the identities of executives tied to that hospital. I link hospitals in this data to characteristics in the American Hospital Association survey and Hospital Compare data. In order to capture the effect of clinical executives, not just their selection into certain types of hospitals, I specifically investigate hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012. This program reduced Medicare payments to hospitals with above average readmission rates, and is plausibly uncorrelated with previously determined leadership teams. I further limit the sample to hospitals with stable executive teams over time, and characterize hospitals as having clinical experience on their team or not.

  I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began, conditional on being penalized. While readmissions directly affects the likelihood of being penalized, mortality rates are an explicit quality measure unrelated to the penalties. Including this as an outcome can help distinguish whether hospitals are improving quality or taking measures to only reduce readmissions.

  I assume that before the penalty, clinical and non-clinical team hospitals behaved similarly along quality dimensions, and that the composition of hospital executive teams determined prior to the penalties is, conditional on being penalized, exogenous. Under these assumptions, I identify the effect of having a clinical executive on response to the penalty. I find that while both types of hospitals, those with clinical executives and those without clinical executives, decrease readmissions after being penalized, non-clinical team hospitals decrease readmissions at a faster rate. The difference equates to approximately 15 additional readmission cases per year for clinical teams, and there is no evidence that this result is driven by differences in readmissions prior to penalties. I find no differential effect of the program on mortality rates.

    This paper contributes to three strands of literature. First, it adds to our understanding of what underlies nonprofit hospital behavior. This literature is extensive, and a review of it is beyond the scope of this paper. To summarize, there are various aspects of nonprofit behavior and objectives for which the evidence of how much it differs from for-profits remains mixed: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}). What underlies this is the need to understand the ways in which nonprofits make choices differently than for-profits. To this point, researchers have found that nonprofits do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}). A paper with a similar methodology to investigate hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likely of shutting down and mix of profitable services. They find that nonprofits and for-profits are equally as likely to shut down after the shock, but only nonprofits adjust their mix of profitable services. Similarly, I investigate a financial shock to hospitals, but I look at the differential impacts of leadership teams within nonprofits. This paper adds to our understanding of inputs into nonprofit hospital objective functions, and whether this is dependent on leadership compositions.

@@

    I merge the leadership information with the AHA survey using the EIN-AHA links I created. From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare data. Hospital Compare contains multiple data sets which I use first to limit to hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), and also for information on my outcome variables, yearly readmission rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.

@@

    \import{Tables}{noMDchg_sumstats.tex}

    In Table \ref{tab:sumstats}, I present summary statistics for various hospital characteristics including which condition hospitals are penalized for, and the readmission and mortality outcome variables. After limiting to penalized hospitals with stable leadership teams, my sample consists of between 302-310 hospitals from 2009-2015, except in the case of AMI outcomes, which is only 225-264 hospitals due to small patient numbers. Fourteen percent of hospitals have someone with clinical experience on their executive team, and the size of the hospitals in the sample varies quite a bit. The average number of beds for a hospital in the sample is 186, only slightly lower than the national average for all nonprofits. Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia. On average, 21\% of patients with one of these conditions would be readmitted. While readmission rates for the specific conditions are similar, the highest likelihood of readmission is for heart failure patients. Alternatively, AMI is the most dangerous condition in terms of mortality rates.

    Figure \ref{fig:weighted_read_mort_graph} shows average readmission rates for hospitals in the sample across time and across the three different conditions plus the weighted average. There are multiple notable points the raw data reveals. First, prior to the penalty, hospitals without clinical executives had consistently higher readmission rates than clinical executive hospitals across all conditions. While there is a gap between clinical team and non-clinical team readmission rates, the lines seem to move mostly parallel. In contrast, 2012 begins a decline in readmission rates over all conditions for all types of hospitals. This makes sense and is consistent with previous literature on HRRP, as hospitals have incentive to reduce readmissions in order to reduce future penalties. However, non-clinical hospital teams decrease readmissions at a faster rate, quickly converging with the rates of clinical team hospitals. While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical leadership teams.

    \begin{figure}[ht!]
        \caption{Readmission rates over time in all penalized hospitals}
        \includegraphics[scale=.55]{Objects/weighted_read_graph.pdf}
        \label{fig:weighted_read_mort_graph}
    \end{figure}

@@

    $$\max\hspace{2mm} \delta R + (1-\delta) u(\theta),$$

@@

    $$\max\hspace{2mm} \delta(L) R(\theta) + (1-\delta(L)) u(\theta)$$

    \noindent where $L$ captures the composition of the leadership team. Therefore, when maximizing this function, the choice of $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs. quality of care.

    Thus, I directly estimate the effect of having clinical experience on readmission rates. However, looking purely at this correlation is a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives. I do two things to address this issue. First, as discussed in detail in Section \ref{sec:data}, I only consider hospitals with stable leadership teams over time. Second, I leverage the HRRP shock, limiting the sample to those that were penalized, which allows me to identify hospitals similar in quality along readmission rates, and to observe changes in behavior after the shock. For the hospitals in my sample, I define their type as having clinical experience on their executive team or not, a time invariant categorization.

@@

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively. As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012. However, I still define $post\_penalty$ as occurring in 2012 or later, the first full year when hospitals would want to respond.

    For identification, I rely on the assumption that, prior to being penalized, clinical and non-clinical team hospitals did not differentially change readmission rates for a reason I do not capture here. I investigate this assumption by estimating event study coefficients, where the coefficients $\beta_j$ now capture the effect of having clinical experience in a particular year, and $\beta_{2011}$ is normalized to zero:

@@

    \subsection{Readmission Rates}

@@

    \import{Tables}{read_results_table.tex}

@@

@@

    \begin{figure}[ht!]
        \caption{Event Study Estimates}
        \includegraphics[scale=.5]{Objects/read_es_graph.pdf}
        \label{fig:wa_eventstudy}
    \end{figure}

    \subsection{Mortality Rates}

    \import{Tables}{mort_results_table.tex}

@@

    In this paper, I study the effect of having clinical experience within hospital leadership on nonprofit hospital behavior. Specifically, I leverage the Hospital Readmissions Reduction Program, which began in 2012, to identify similar quality hospitals along the dimension of readmission rates. For these similar hospitals, I estimate the difference in readmissions after being penalized between hospitals with clinical executives and hospitals without clinical executives. I find that while the penalty program led both types of hospitals to decrease readmissions after being penalized, those without clinical experience on their executive decreased rates more quickly than non-clinical team hospitals.

    This finding reveals important information about hospital objectives, specifically that the composition of a hospital's leadership team affects the value placed on profit (or in this case, avoiding loss) vs. quality of care. Therefore, not-for-profit vs. for-profit status should not be the only characteristic considered by policymakers when designing incentives to increase quality of care.

@@

	% \appendix

 %    \section{Data}\label{appendixdata}

 %    \subsection{Gathering Hospital Leadership Names}

 %    There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper. However, using the NonProfit Explorer API seems to be the most straightforward. At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/nonprofits/api}{https://projects.propublica.org/nonprofits/api}.

 %    There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs. The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code. I query only nonprofits categorized as E20 (hospitals), E21 (community health systems), and E22 (general hospitals). The API has a pagination limit of 100, meaning I can only pull information on 100 hospitals at a time. Therefore, I filter the query further to only consider one state at a time. The only state that has more than 100 entities registered is California, and thus I subset the California query even further by names that include the word "hospital" and names that don't. I combine all of these subsets and have information on each nonprofits Employee Identification Number. There are 5,588 EINs total in this list. This acts as a list of entities for which I can pull more information.

 %    I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN. I save the name, secondary name, state, and zip code, all of which do not vary by year. I also save each year's URL that links to the Tax Form 990 PDF. For the sake of a comprehensive data set, I keep years 2006-2020 (I later limit to 2009-2016 when focusing on the Hospital Readmissions Reduction Program). Thus, I finish this step with a panel data set of EIN characteristics and PDF locations. Importantly, there are multiple types of Tax Form 990s depending on the size of the nonprofit. In many cases, one nonprofit has at least two different forms filed in a given year. I filter out any EIN-years for which there are no PDFs on file. The data on PDF locations contains 4,012 EINs and 61,363 EIN-year-tax forms.

 %    It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes). Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location. First, I will discuss limitations and cleaning that I do to the AHA data and tax form data separately before doing any matching.

 %    I download the AHA data from Wharton Research Data Services. I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as nonprofit or state/community, and those that are general acute care. I also filter out any hospitals who weren't present in the data (or change system ID) in 2009-2015, meaning they either closed or were acquired. Due to the survey nature of this data, a hospital name may look slightly different from one year to the next. For example, "Waldo County General Hospital" is also Waldo County General Hospital Maine Health". Further, zip codes may change by one or two digits, making them unreliable to match based on. To deal with this, I first keep only unique AHA ID, name, zip, state, and system name combinations. Then, I convert the data from long to wide so that each AHA ID occurs only once, but may have multiple names, zip codes, or system names associated with it.

 %    I consider which nonprofit entities are not likely to be hospitals and drop them. There are numerous foundations or auxiliary nonprofits with the purpose of raising funds for the hospital, but do not actually care for patient. I filter out any nonprofit with "foundation" or "auxiliary" in the name. I also filter out various specialty centers that fell into the general hospital category, such as hospice or cancer centers.

 %    I then proceed matching based on names in multiple layers. I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc. Next, I take each AHA name and look for exact matches in a nonprofit's first or secondary name only for nonprofits in the same state as the AHA hospital. When an exact match is found, I record the link between AHA ID and EIN. In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of AHA nonprofit hospitals in the sample.

 %    In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc. That way if there are subtle differences in names, removing common words may allow for an exact match. Again, I take each AHA hospital name and look for exact matches in the nonprofits within the same state. This adds an additional 90 hospital matches, accounting for a total of 34.5\% of AHA hospitals.

 %    In some cases, tax forms are associated with a system of hospitals instead of one individual hospital. Thus, I create another variable that captures the names of systems which match. The process of matching is the same, the only difference is that now I am considering system names from the AHA data instead of hospital names. A total of 1,136 AHA hospitals have an EIN match on hospital name, system name, or both. This is roughly 41\% of the sample. These hospitals become the main sample for my analysis. While not capturing a large majority of AHA hospitals that are categorized as nonprofit, I would rather be conservative in selecting hospitals which I am confident that the characteristics and measurement of the independent and dependent variables are accurate.

 %    (add a paragraph about validity of matches)

>>>>

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{papercitations.bib}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}

@@

	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Programs}

@@

    \begin{abstract}
		{\small
        Nonprofit firms have diverse goals and missions that make understanding their true objectives and behaviors difficult. A setting where this is particularly relevant is that of nonprofit hospitals, which make up the majority of hospitals in the US, yet there is mixed evidence on what drives their behavior. In some ways they seem to act like for-profits, while in other ways they may care more about prestige or quantity. In this paper, I investigate whether the leadership team composition of nonprofit hospitals is correlated with more for-profit behavior. I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether having clinical experience on a hospital executive team affects hospital behavior. In particular, I leverage several pay-for-performance initiatives in the US in 2012, and estimate the difference in response between for-profits, nonprofits, and clinical nonprofits. I find that while readmissions decreased on average for all types of hospitals, non-clinical executive team hospital decreased readmissions at a faster rate than those with a clinical executive team, and behaved more similarly to for-profit hospitals. The difference equates to approximately 15 additional readmissions for clinical team hospitals.

@@

  The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms. Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing mission-based goals. Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}). An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (\cite{sloan2000not}). In this paper, I leverage a large-scale shock to hospitals in the US to investigate whether differences in leadership composition reveal characteristics of hospital objective functions, namely the weight placed on profit vs. utility gained from benefiting society.

  Hospitals in particular have been a common focus when seeking to understand nonprofit firms. This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits. Second, nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}). Thus, nonprofit behavior in the hospital industry is relevant and directly affects consumers of health care. However, it remains unclear what type of objective function drives nonprofit hospitals (\cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}). In particular, the extent to which nonprofit hospitals gain utility from bettering society (e.g. increased charity care, high quality of care, etc.) is unknown.

  What has not been considered is how the leadership team within nonprofits potentially drives objectives. In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than for-profit or not-for-profit status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}). Within nonprofit hospitals, there is variation in the propensity to hire executives with a clinical background, which could reasonably change how much weight the hospital places on profit vs. quality of care when making decisions. I contribute to our understanding of hospital and nonprofit behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.

  Using publicly available tax forms, I construct a novel data set of nonprofit hospital executives from 2009-2015. I link hospitals in this data to characteristics in the American Hospital Association survey and Hospital Compare data. In order to capture the effect of clinical executives, not just their selection into certain types of hospitals, I consider the sample of hospitals whose leadership team remains stable over time, and characterize them as having clinical executive experience on their team or not, a time-invariant categorization. I then leverage the Hospital Readmissions Reduction Program (HRRP) of 2012. This program reduced Medicare payments to hospitals with above average readmission rates, and is a shock that is plausibly uncorrelated with previously determined leadership teams.

  I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began. While readmissions directly affects the likelihood of being penalized, mortality rates are an explicit quality measure unrelated to the penalties. Including this as an outcome can help distinguish whether hospitals are improving quality or only taking measures to reduce readmissions. Not all hospitals were penalized, but all hospitals (with few exceptions) could have been penalized if readmission rates were high enough. Thus, I present estimates first for all hospitals. However, these estimates are biased if I am simply capturing the effect of clinical experience on already being a good quality hospital. Thus, I present two further specifications: one limiting to only hospitals who were eventually penalized, and another where I limit the sample further to hospitals which have an observably similar match.

  I assume that before the penalty, clinical and non-clinical team hospitals behaved similarly along quality dimensions, and that the composition of hospital executive teams determined prior to the penalties is exogenous. Under these assumptions, I identify the effect of having a clinical executive on hospital response to the penalty. I find that while both types of hospitals, those with clinical executives and those without clinical executives, decrease readmissions after being penalized, non-clinical team hospitals decrease readmissions at a faster rate. The difference equates to approximately 15 additional readmission cases per year for clinical teams, and there is no evidence that this result is driven by differences in readmissions prior to penalties. However, I find no differential effect of the program on mortality rates.

    This paper contributes to three strands of literature. First, it adds to our understanding of what underlies nonprofit hospital behavior. This literature is extensive, and a review of it is beyond the scope of this paper. To summarize, there are various aspects of nonprofit behavior and objectives for which the evidence of how much it differs from for-profits remains mixed: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}). What underlies this is the need to understand the ways in which nonprofits make choices differently than for-profits. To this point, researchers have found that nonprofits do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}). A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likely of shutting down and mix of profitable services. They find that nonprofits and for-profits are equally as likely to shut down after the shock, but only nonprofits adjust their mix of profitable services. Similarly, I investigate a financial shock to hospitals, but I look at the differential impacts of leadership teams within nonprofits. This paper adds to our understanding of inputs into nonprofit hospital objective functions, and whether this is dependent on leadership compositions.

@@

    I merge the leadership information with the AHA survey using the EIN-AHA links I created. From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare data. Hospital Compare contains multiple data sets which I use first to limit to hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), and also for information on outcome variables: readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.

@@

    \subsection{Summary Statistics}

    \import{Tables}{overall_sumstats.tex}

    \import{Tables}{sample_sumstats.tex}

    In Table \ref{tab1:sumstats}, I present summary statistics for various hospital characteristics including which condition hospitals are penalized for, and the readmission and mortality outcome variables. After limiting to penalized hospitals with stable leadership teams, my sample consists of between 302-310 hospitals from 2009-2015, except in the case of AMI outcomes, which is only 225-264 hospitals due to small patient numbers. Fourteen percent of hospitals have someone with clinical experience on their executive team, and the size of the hospitals in the sample varies quite a bit. The average number of beds for a hospital in the sample is 186, only slightly lower than the national average for all nonprofits. Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia. On average, 21\% of patients with one of these conditions would be readmitted. While readmission rates for the specific conditions are similar, the highest likelihood of readmission is for heart failure patients. Alternatively, AMI is the most dangerous condition in terms of mortality rates.

    Figure \ref{fig:weighted_read_mort_graph} shows average readmission and mortality rates across time and across the three different conditions, along with the weighted average. There are multiple notable points the raw data reveals. First, prior to the penalty, hospitals without clinical executives had consistently higher readmission rates than clinical executive hospitals across all conditions. While there is a gap between clinical team and non-clinical team readmission rates, the lines seem to move mostly parallel. In contrast, 2012 begins a decline in readmission rates over all conditions for all types of hospitals. This makes sense and is consistent with previous literature on HRRP, as hospitals have incentive to reduce readmissions in order to reduce future penalties. However, non-clinical hospital teams decrease readmissions at a faster rate, quickly converging with the rates of clinical team hospitals. While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical leadership teams. On the right panel, average mortality rates over time reveal no clear change due to the penalty, and no consistent pattern in the difference between clinical and non-clinical hospital teams.

    \begin{figure}[ht!]
        \caption{Outcomes over time in all penalized hospitals}
        \includegraphics[scale=.5]{Objects/weighted_read_mort_graph.pdf}
        \label{fig:weighted_read_mort_graph}
    \end{figure}

    \import{Tables}{hosp_correlations.tex}

@@

    $$\max_{\theta}\hspace{2mm} \delta R + (1-\delta) u(\theta),$$

@@

    $$\max_{\theta}\hspace{2mm} \delta(L) R(\theta) + (1-\delta(L)) u(\theta)$$

    \noindent where $L$ captures the composition of the leadership team. Therefore, when maximizing this function, the choice of $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs. quality of care:

    $$\theta = f(\delta(L)).$$

    Thus, I directly estimate the effect of having clinical experience on readmission rates. Additionally, I estimate the effect on mortality rates to capture any spillover effects of improving readmissions into another dimension of quality. However, looking purely at this correlation is a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives. I do two things to address this issue. First, as discussed in detail in Section \ref{sec:data}, I only consider hospitals with stable leadership teams over time. Second, I leverage the HRRP shock, limiting the sample to those that were penalized, which allows me to identify hospitals similar in quality, and to observe changes in behavior after the shock. For the hospitals in my sample, I define their type as having clinical experience on their executive team or not, a time invariant categorization.

@@

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively. As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012. However, I still define $post\_penalty$ as occurring in 2012 or later, the first full year when hospitals could respond.

    For identification, I rely on the assumption that, prior to being penalized, clinical and non-clinical team hospitals did not differentially change readmission or mortality rates for a reason I do not capture here. I investigate this assumption by estimating event study coefficients, where the coefficients $\beta_j$ now capture the effect of having clinical experience in a particular year, and $\beta_{2011}$ is normalized to zero:

@@

    \import{Tables}{MD_noMD_readmort_synthdid.tex}

    \import{Tables}{forprofit_synth.tex}

    \import{Tables}{MD_noMD_uncompCMI_synth.tex}

    \import{Tables}{forprofit_uncompCMI_synth.tex}

@@

@@

    \begin{figure}[ht!]
        \caption{Event Study Estimates, Readmission}
        \includegraphics[scale=.5]{Objects/read_es_graph.pdf}
        \label{fig:wa_eventstudy}
    \end{figure}

@@

    Estimates for the coefficient of interest from Equation \ref{eq:main} with mortality rates are presented in Table \ref{tab:mort_results}. Column 1 presents my estimate of the effect of having a clinical experience team on the weighted average of mortality. Columns 2, 3, and 4 present estimates where the sample is limited to hospitals who are penalized for the outcome being considered. The coefficients are all positive, ranging from .13 to .31, but none of them are statistically significant. This is consistent with prior literature that finds hospitals do not change mortality rates after HRRP, and this result reveals that holds across different leadership teams.

    I investigate whether mortality rates seem to be driven by a pre-trend, or whether the effect of clinical experience on mortality rates differs across time after the penalty occurs. Event study estimates from Equation \ref{eq:es} for the weighted average mortality model is shown in Figure \ref{fig:wa_eventstudy_mort}. These estimates reveal that hospital teams with clinical experience do not differentially change mortality rates compared to non-clinical experience teams before or after the penalties begin.

    \begin{figure}[ht!]
        \caption{Event Study Estimates, Mortality}
        \includegraphics[scale=.5]{Objects/mort_es_graph.pdf}
        \label{fig:wa_eventstudy_mort}
    \end{figure}

@@

    In this paper, I study the effect of having clinical experience within hospital leadership on nonprofit hospital behavior. Specifically, I leverage the Hospital Readmissions Reduction Program, which began in 2012, to identify similar quality hospitals along the dimension of readmission rates. For these similar hospitals, I estimate the difference in readmissions after being penalized between hospitals with clinical executives and hospitals without clinical executives. I find that while the penalty program led both types of hospitals to decrease readmissions after being penalized, those without clinical experience on their executive team decreased rates more quickly than non-clinical team hospitals.

    This finding reveals important information about hospital objectives, specifically that the composition of a hospital's leadership team affects the value placed on profit (or in this case, avoiding loss) vs. quality of care. Therefore, policymakers should take into consideration the composition of the leadership team when designing pay-for-performance incentives.

@@

\appendix

 \section{Data}\label{appendixdata}

\subsection{Gathering Hospital Leadership Names}

There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper. However, using the NonProfit Explorer API seems to be the most straightforward. At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/nonprofits/api}{https://projects.propublica.org/nonprofits/api}.

There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs. The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code. I query only nonprofits categorized as E20 (hospitals), E21 (community health systems), and E22 (general hospitals). The API has a pagination limit of 100, meaning I can only pull information on 100 hospitals at a time. Therefore, I filter the query further to only consider one state at a time. The only state that has more than 100 entities registered is California, and thus I subset the California query even further by names that include the word "hospital" and names that don't. I combine all of these subsets and have information on each nonprofits Employee Identification Number. There are 5,588 EINs total in this list. This acts as a list of entities for which I can pull more information.

I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN. I save the name, secondary name, state, and zip code, all of which do not vary by year. I also save each year's URL that links to the Tax Form 990 PDF. For the sake of a comprehensive data set, I keep years 2006-2020 (I later limit to 2009-2016 when focusing on the Hospital Readmissions Reduction Program). Thus, I finish this step with a panel data set of EIN characteristics and PDF locations. Importantly, there are multiple types of Tax Form 990s depending on the size of the nonprofit. In many cases, one nonprofit has at least two different forms filed in a given year. I filter out any EIN-years for which there are no PDFs on file. The data on PDF locations contains 4,012 EINs and 61,363 EIN-year-tax forms.

It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes). Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location. First, I will discuss limitations and cleaning that I do to the AHA data and tax form data separately before doing any matching.

I download the AHA data from Wharton Research Data Services. I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as nonprofit or state/community, and those that are general acute care. I also filter out any hospitals who weren't present in the data (or change system ID) in 2009-2015, meaning they either closed or were acquired. Due to the survey nature of this data, a hospital name may look slightly different from one year to the next. For example, "Waldo County General Hospital" is also Waldo County General Hospital Maine Health". Further, zip codes may change by one or two digits, making them unreliable to match based on. To deal with this, I first keep only unique AHA ID, name, zip, state, and system name combinations. Then, I convert the data from long to wide so that each AHA ID occurs only once, but may have multiple names, zip codes, or system names associated with it.

I consider which nonprofit entities are not likely to be hospitals and drop them. There are numerous foundations or auxiliary nonprofits with the purpose of raising funds for the hospital, but do not actually care for patient. I filter out any nonprofit with "foundation" or "auxiliary" in the name. I also filter out various specialty centers that fell into the general hospital category, such as hospice or cancer centers.

I then proceed matching based on names in multiple layers. I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc. Next, I take each AHA name and look for exact matches in a nonprofit's first or secondary name only for nonprofits in the same state as the AHA hospital. When an exact match is found, I record the link between AHA ID and EIN. In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of AHA nonprofit hospitals in the sample.

In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc. That way if there are subtle differences in names, removing common words may allow for an exact match. Again, I take each AHA hospital name and look for exact matches in the nonprofits within the same state. This adds an additional 90 hospital matches, accounting for a total of 34.5\% of AHA hospitals.

In some cases, tax forms are associated with a system of hospitals instead of one individual hospital. Thus, I create another variable that captures the names of systems which match. The process of matching is the same, the only difference is that now I am considering system names from the AHA data instead of hospital names. A total of 1,136 AHA hospitals have an EIN match on hospital name, system name, or both. This is roughly 41\% of the sample. These hospitals become the main sample for my analysis. While not capturing a large majority of AHA hospitals that are categorized as nonprofit, I would rather be conservative in selecting hospitals which I am confident that the characteristics and measurement of the independent and dependent variables are accurate.

(add a paragraph about validity of matches)

 Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs. In the data set of hospital PDF URLs that I collected earlier, I limit to the hospitals with solid matches described above. I then loop through each EIN, downloading each PDF locally and using tesseract package in R to extract text from the relevant pages of the PDF using OCR text extraction methods. In particular, I loop through each page of the PDF, look for the title associated with leadership names: ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees", and save all the text from any pages where this title is found. I save the text to a list of all EIN, years present.
